Biostat III exercises in R
===========
Laboratory exercise 1b
-----------

### Suggested solutions by
Author: Annika Tillander, 2014-01-30  
Edited: Andreas Karlsson, 2015-02-27

-----------

```{r setup, cache=FALSE, message=FALSE, echo=FALSE}	
library('knitr')	
read_chunk('q1.R')
opts_chunk$set(cache=FALSE)
```

You may have to install the required packages the first time you use them. You can install a package by `install.packages("package_of_interest")` for each package you require.

```{r loadDependecies, message=FALSE}
```

Number of events and number lost summarised by year.

```{r eventsPerMonth}
```

Following are the life table estimates. Note that in the lectures, when we estimated all-cause survival, there were 8 deaths in the first interval. One of these died of a cause other than cancer so in the cause-specific survival analysis we see that there are 7 ‘deaths’ and 1 censoring (lifetab uses the term ‘nlost’ for number lost to follow-up) in the first interval.

```{r lifeTable}
```

Following is a table of Kaplan-Meier estimates. It is not clear from the table, if the person censored at time 2 was at risk when the other person dies at time 2. Following the table is a graph of the survival function.


```{r KaplanMeier, message=FALSE}
```

Kaplan-Meier esimates by sex with a table of the number at risk.

```{r KaplanMeierNumberAtRisk, message=FALSE, fig.width=7, fig.height=7}
```
